### 2022-06-09 README.Rmd
Small changes + LICENSE change
Small changes in code and changed license from MIT to GPL2 since the package depend on INLA which is GPL2

```{r global_options, include = FALSE}
library(knitr)
options(width = 120)
opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'Figs/',
               include = TRUE, warning = FALSE, message = FALSE)
```

# Bremla
This Repository contains the Bremla package for uncertainty quantification and synchronization for layer-counted proxy archives, including the data and code to reproduce the results for

Myrvoll-Nilsen, E., Riechers, K., Rypdal, M. & Boers, N. (2022). Comprehensive uncertainty estimation of the timing of Greenland warmings of the Greenland Ice core records. Climate of the Past, x(x),x-x. doi:xxxxx

and

Myrvoll-Nilsen, E., Riechers, K. & Boers, N. (202x). Tie-point paper (title TBD)

## installation

The simplest way to install the package is to install the remotes package and run
```{r github-install, eval=FALSE}
#install.packages("remotes")
remotes::install_github("eirikmn/bremla")
```

To get started, see the documentation '?bremla' and the associated example.

The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [iceandclimate.nbi.ku.dk](Centre for Ice and Climate) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [https://www.sciencedirect.com/science/article/pii/S0277379114003485](Rasmussen et al. (2014)). Also includes the tie-point presented in [https://cp.copernicus.org/articles/14/1755/2018/](Adolphi et al. (2018)) and [https://www.cambridge.org/core/journals/radiocarbon/article/testing-and-improving-the-intcal20-calibration-curve-with-independent-records/D72D9214C47FE9441B5E730D33DCCE3D](Muscheler et al. (2020)). Other data and tie-points should work as well.

Warning: The provided real data example generates several thousand simulated chronologies with individual lengths exceeding 18,000. More if synchronized chronologies is also computed. This will require a significant amount of memory. If needed, reduce the number of samples generated or free up memory before use.

## Example
This is a real data example which produces 5000 simulations from a synchronized time scale.
```{r example, message=FALSE, warning=FALSE, include=FALSE}
require(INLA)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O

data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
formula = dy~-1+depth2

events=list(locations = events_rasmussen$depth,
             locations_unit="depth",degree=1)
control.transition_dating=list(age.ref=age.reference,age.label="years (yb2k)")
object = bremla(formula,data,reference.label="GICC05",
                  nsims=5000,
                  events=events,
                  synchronization=list(method="adolphi"),
                  control.fit=list(method="inla"),
                  control.sim=list(synchronized=TRUE),
                  print.progress=TRUE )
summary(object)
plot(object)
```

## Attribution

This code is associated and written for the papers Myrvoll-Nilsen et al. (2022) and Myrvoll-Nilsen et al. (202x) mentioned above. Feel free to use the code, but please cite the accompanying papers.

## License

The code in this repository is made available under the terms of the MIT License. For details, see LICENSE file.