### 2023-07-05 inst/reproduce_results/test.Rmd
Fixed some bugs and changed the hyperparameter posterior sampler
no samples hyperparameters from joint distribution

---
output: github_document
---

<!-- results.md is generated from results.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/reproduce/results/results-",
  dev = "png",
  out.width = "100%"
)
```

# Extract DO events from Corrick pdf

```{r initialize, eval=TRUE}
library(pdftools)
library(data.table)
library(dplyr)
library(stringr)

# ttt = pdf_text("inst/reproduce_results/corricktable.pdf")
# 
# data = toString(ttt)
# data <- read.table(text = data, sep = "\n", as.is = TRUE)
# data <- data[-c(1, 2, 3, 4, 5, 6, 7, 14, 20, 26, 34, 47, 52, 58, 65, 66, 67), ]
# data <- gsub("[,]", "", data)
# data <- gsub("[$]", "", data)
# data <- gsub("\\s+", ",", gsub("^\\s+|\\s+$", "",data))
# 
# 
# 
# df <- ttt %>% 
#   read_lines() %>%    # separate lines
#   grep('^\\s{2}\\w', ., value = TRUE) %>%    # select lines with states, which start with space, space, letter
#   paste(collapse = '\n') %>%    # recombine
#   read_fwf(fwf_empty(.)) %>%    # read as fixed-width file
#   mutate_at(-1, parse_number) %>%    # make numbers numbers
#   mutate(X1 = sub('*', '', X1, fixed = TRUE))    # get rid of asterisks in state names
# 
# df

colnames = c("age", "avg error", "negative error"," positive error",
             "precision", "country", "region")
eventnames = c("GI 1e", "GI 3", "GI 4", "GI 5.1", "GI 5.2", "GI 6", "GI 7a",
               "GI 7c", "GI 8c", "GI 9", "GI 10", "GI 11","GI12c", "GI 13a", 
               "GI 13c", "GI14c", "GI 15.1", "GI 15.2", "GI 16.1a", "GI 16.1c",
               "GI 16.2", "GI 17.1a", "GI 17.1c", "GI 17.2", "GI 18", "GI 19.2",
               "GI 20c", "GI 21.1a", "GI 21.1e", "GI 22g", "GI 23.2", "GI 24.1a",
               "GI 24.1c", "GI 24.2", "GI 25c")

fullfile = scan("corrictabletext.txt", what="character",sep="\t")
findDOlines = gregexpr("GI",fullfile)
DOlines = numeric(length(eventnames))
teller=0
for(i in 1:length(findDOlines)){
  if(findDOlines[[i]][1] > -1){
    teller = teller+1
    DOlines[teller] = i
  }
}
DOlines = c(DOlines, length(fullfile)+1)

dflist = list()

#for(i in 19:19){
for(i in 2:length(DOlines)){  
  startline = DOlines[i-1]+1
  endline = DOlines[i]-1
  eventname = eventnames[i-1]
  rownames=c()
  rows = c()
  for(iterline in startline:endline){
    currentline = fullfile[iterline]
    elements = str_split(currentline," ")[[1]]
    elemarray = c()
    for(elem in 1:length(elements)){
      if(elements[elem] != ""){
        elemarray=c(elemarray, elements[elem])
      }
    }
    rownames = c(rownames, elemarray[1])
    rows=rbind(rows, elemarray[-1])
  }
  colnames(rows) = colnames
  rownames(rows) = rownames
  dflist[[eventname]] = rows
}

dflist[[1]]

```
```{r functions}

cleanstring = function(str,do.tolower=TRUE){
  str = gsub("\\s+|\\.|,|_|-", "", str, perl = TRUE)

  if(do.tolower){
    return(tolower(str))
  }else{
    return(str)
  }
}

stringcut = function(str,expr="_",before=TRUE,margin=1){
  pos <- regexpr(paste0("\\",expr), str)
  if(pos==-1){
    return(str)
  }else{
    if(before){
      return(substr(str,1,pos-margin))
    }else{
      return(substr(str,pos+margin,nchar(str)))
    }
  }
}
which.col = function(mat, val){
  loc=c()
  for(i in 1:ncol(mat)){
    ran = range(mat[!is.na(mat[,i]),i])
    if(ran[1] <= val && ran[2] >= val){
      loc=c(loc, i)
    }
  }
  if(length(loc)==0){
    warning("Could not find element in matrix columns!\n")
  }else if(length(loc)>1){
    warning("Element found in multiple columns...\n")
  }
  return(loc)
}
```

```{r dflist_bygroup}

codenames = c()
for(i in 1:length(dflist)){
  codenames = c(codenames, rownames(dflist[[i]]) )
}
codenames = unique(codenames)

records = c()

for(i in 1:length(codenames)){
  locstr = stringcut(codenames[i],expr="_",before=FALSE)
  groupstr = stringcut(codenames[i],expr="_",before=TRUE)
  groupstr = stringcut(groupstr,expr="(",before=TRUE)

  records = c(records, paste0(groupstr,"_",locstr))
}
records = unique(records)
nrecords = length(records)

fullfile = scan("corrictabletext.txt", what="character",sep="\t")

findDOlines = gregexpr("GI",fullfile)
DOlines = numeric(length(eventnames))

teller=0
for(i in 1:length(findDOlines)){
  if(findDOlines[[i]][1] > -1){
    teller = teller+1
    DOlines[teller] = i
  }
}
DOlines = c(DOlines, length(fullfile)+1)

#for(i in 19:19){
#for(i in 1:nrecords){  
  #findeventlines = gregexpr(tolower(records[i]),tolower(fullfile))
endline = length(fullfile)
df2list = list()
for(iterline in 4:endline){
  currentline = fullfile[iterline]
  elements = str_split(currentline," ")[[1]]

  elemarray = c()
  for(elem in 1:length(elements)){
    if(elements[elem] != ""){
      elemarray=c(elemarray, elements[elem])
    }
  }
  if("GI" %in% elemarray[1]){
  }else{
    tempdf = data.frame(age=elemarray[2], avgerror=elemarray[3], negerror=elemarray[4],
                      poserror=elemarray[5],precision=elemarray[6])

    eventind = max(which(DOlines<iterline))
    eventstr = eventnames[eventind]
    rownames(tempdf) = eventstr

    namestr = elemarray[1]
    groupstr = stringcut(namestr,"_",before=TRUE)
    groupstr = stringcut(groupstr,"(",before=TRUE)
    locstr = stringcut(namestr,"_",before=FALSE)
    recstring = paste0(groupstr,"_",locstr)

    df2list[[recstring]] = rbind(df2list[[recstring]],tempdf)

  }
  #}

}

df2list[[1]]

```

`
```{r rgeneric-linramp}
rgeneric.uneven.AR1 = function( #specifies necessary functions for INLA to define the linear ramp model
  cmd = c("graph", "Q","mu", "initial", "log.norm.const", "log.prior", "quit"),
  theta = NULL)
{

  envir = environment(sys.call()[[1]])

  linramp = function(t,t0=0,dt=1,y0=0,dy=1){ #linear ramp function
    y = numeric(length(t))
    y = y0 + dy*(t-t0)/dt

    y[t<t0]=y0
    y[t>t0+dt]=y0+dy

    return(y)
  }

  interpret.theta = function() { #helpful function to transform back from internal parametrization
    y0 = theta[3]
    dy = (theta[4])
    y1 = y0+dy
    t0 = theta[1]
    Dt = exp(theta[2])
    t1=t0+Dt

    prec = exp(theta[6])
    tau = exp(theta[5])
    return(list(y0=y0,dy=dy,y1=y1,t0=t0,t1=t1,prec=prec,tau=tau,Dt=Dt))
  }

  mu = function() { #mean vector defined as a linear ramp
    if(!is.null(envir)){
      timepoints=get("timepoints",envir)

    }
    param = interpret.theta()
    y0=param$y0; dy=param$dy; y1=param$y1; t0=param$t0; t1=param$t1; Dt = param$Dt; dt=Dt

    mvek = linramp(timepoints,t0=t0,dt=dt,y0=y0,dy=dy)

    return(mvek)
  }

  graph = function(){ #graphs of conditional dependence structure. 1 where Q[i,j] != 0, 0 where Q[i,j] = 0
    G = Q()
    G[G != 0] = 1
    return (G)
  }

  Q = function(){ #inverse covariance matrix: AR(1) that allow for unequal spacing
    if(!is.null(envir)){
      n=get("n",envir)
      timepoints=get("timepoints",envir)
    }
    param=interpret.theta()
    ii = 1:n
    jj = 1:n

    rhos=rep(NA,n)
    rhos[2:n] = exp(-diff(timepoints)/param$tau)
    kappa0 = param$prec

    xx = rep(NA,2*n-1)
    xx[1] = 1+rhos[2]^2/(1-rhos[2]^2)
    xx[2] = 1/(1-rhos[n]^2)
    xx[3:n] = 1/(1-rhos[2:(n-1)]^2) + rhos[3:n]^2/(1-rhos[3:n]^2)
    xx[(n+1):(2*n-1)] = -rhos[2:n]/(1-rhos[2:n]^2)

    xx = kappa0*xx

    i = c(1L, n, 2L:(n - 1L), 1L:(n - 1L))
    j = c(1L, n, 2L:(n - 1L), 2L:n)

    Q = Matrix::sparseMatrix(
      i = i,
      j = j,
      x = xx,
      symmetric = TRUE
    )
    #diag(Q)=diag(Q)

    return (Q)
  }

  log.norm.const = function(){ #INLA computes this automatically
    return(numeric(0))
  }

  log.prior = function(){
    if(!is.null(envir)){
      tslutt=get("tslutt",envir)
      tstart=get("tstart",envir)
      ystart=get("ystart",envir)
      #log.theta.prior=get("log.theta.prior",envir)
      # rescale.y=get("rescale.y",envir)
    }

    #if(!is.null(log.theta.prior)){
    #  lprior = log.theta.prior(theta)
    #}else{
      params = interpret.theta()

      #log-priors are given for internal parametrisation using the change of variables theorem

      lprior = dnorm(theta[1],mean=round(0.5*(tslutt+tstart)),sd=50,log=TRUE) #t0
      lprior = lprior + dgamma(exp(theta[2]),shape=1.0,rate=0.02,log=TRUE) + theta[2] #dt
      # if(rescale.y){
      #   lprior = lprior + dnorm(theta[3],mean=ystart,sd=0.025,log=TRUE) #y0
      #   lprior = lprior + dnorm(theta[4],mean=1-ystart,sd=0.01,log=TRUE) #dy
      #   lprior = lprior + dgamma(exp(theta[5]),1,rate = 3,log=TRUE) + theta[6] #tau/rho
      #   lprior = lprior + dgamma(exp(theta[6]),0.1,rate = 1,log=TRUE) + theta[6] #sigma/kappa
      # }else{
      lprior = lprior + dnorm(theta[3],mean=ystart,sd=5,log=TRUE) #y0
      lprior = lprior + dnorm(theta[4],mean=0,sd=10.0,log=TRUE) #dy
      lprior = lprior + dgamma(exp(theta[5]),2.5,rate = 0.15,log=TRUE) + theta[6] #tau/rho
      lprior = lprior + dgamma(exp(theta[6]),2,rate = 0.15,log=TRUE) + theta[6] #sigma/kappa
      # }
    #}

    return (lprior)
  }

  initial = function(){
    ini = c(0,0,0,0,1,2)
    return (ini)
  }

  quit = function(){
    return ()
  }
  if (!length(theta)) theta = initial()

  cmd = match.arg(cmd)
  val = do.call(cmd, args = list())
  return (val)
}
```

```{r linrampfunction}

linrampfunc = function(x,y, steplength=0.005,nsims=0, opt.params=NULL,imp.fit=TRUE, rescale.y.factor=1, print.progress=TRUE){

  n = length(x)

  input_x=x
  input_y=y
  if(print.progress) cat("Initializing linear ramp fit using INLA.\n",sep="")
  df_event = data.frame(xx=x, yy=y)

    #timepoints = input_x
  timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)

    ## default initial values for optim function
  if(is.null(opt.params)){
    optparams = c(round(length(timepoints)/2),round(length(timepoints)/10),df_event$yy[1],df_event$yy[length(timepoints)]-df_event$yy[1])
  }else{
    defoptparams = c(round(length(timepoints)/2),round(length(timepoints)/10),df_event$yy[1],df_event$yy[length(timepoints)]-df_event$yy[1])
    optparams = control.linramp$opt.params
    for(i in 1:length(defoptparams)){
      if(is.na(optparams[i])){
        optparams[i] = defoptparams[i]
      }
    }
  }

  df0=data.frame(y=df_event$yy*rescale.y.factor,
                 x=df_event$xx)

  t_start = df0$time[1];t_end = df0$time[n];y_start = df0$y[1]
  #library(numDeriv)
  #library(compiler)
  y = df0$y
  if(print.progress) cat("Using 'optim' to find initial positions for hyperparameters in INLA.\n",sep="")
  ## for stability the x-axis is transformed to values ranging from 1:n.

  df0$time = timepoints

  ## Finding initial values in INLA optimization procedure by first using a simple optimization
  ## gradient and cost function given here

  if(!is.null(imp.fit)){
    minfun.grad = function(param, args = NULL){
      return (grad(minfun, param, args=args, method.args = list(r=6)))
    }
    minfun = function(param, args = NULL){
      yhat = linramp(timepoints,t0=param[1],dt=param[2],y0=param[3],dy=param[4])
      mse = sum((args$y-yhat)^2)
      return(sqrt(mse))
    }

    ## perform optimization to find good starting values for INLA
    param=optparams
    args=list(y=y)
    fit = optim(param,
                fn = minfun,
                gr = minfun.grad,
                method = "BFGS",
                control = list(
                  abstol = 0,
                  maxit = 100000,
                  reltol = 1e-11),
                args = args)

    ### use least squares estimates for fixed effects as initial values in inla

    muvek = linramp(timepoints,t0=fit$par[1],dt=fit$par[2],y0=fit$par[3],dy=fit$par[4])
    #plot(timepoints,y); lines(timepoints,muvek)
    init = c(fit$par[1],log(fit$par[2]),fit$par[3],fit$par[4],0,0)
  }

  if(print.progress) cat("Fitting linear ramp model in INLA using rgeneric model specification...\n",sep="")
  ## creating linear ramp INLA model using rgeneric framework. Requires further specification, see "rgeneric.uneven" function
  time.startinla = Sys.time()
  model.rgeneric = INLA::inla.rgeneric.define(rgeneric.uneven.AR1,n=n,
                                              tstart=timepoints[1],
                                              tslutt=timepoints[n],
                                              ystart=y[1],
                                              timepoints = timepoints)#,
                                              #log.theta.prior=control.linramp$log.theta.prior)
  formula = y ~ -1+ f(idx, model=model.rgeneric)

  if(!is.null(imp.fit)){
    r = INLA::inla(formula,family="gaussian", data=data.frame(y=df0$y,idx=1:n),#as.integer(df0$time)),
                   control.mode=list(theta=init,
                                     restart=TRUE),
                   num.threads = 1,
                   verbose=FALSE,
                   silent=FALSE,
                   control.inla=list(h=steplength),
                   control.family = list(hyper = list(prec = list(initial = 12, fixed=TRUE))) )#, num.threads = 1)

  }else{
    r = INLA::inla(formula,family="gaussian", data=data.frame(y=df0$y,idx=1:n),#as.integer(df0$time)),
                   control.mode=list(restart=TRUE),
                   num.threads = 1,
                   verbose=FALSE,
                   silent=FALSE,
                   control.inla=list(h=steplength),
                   control.family = list(hyper = list(prec = list(initial = 12, fixed=TRUE))) )#, num.threads = 1)

  }

  time.endinla = Sys.time()
  elapsedinla = difftime(time.endinla,time.startinla,units="secs")[[1]]
  if(print.progress) cat("Completed in ",elapsedinla," seconds.\n",sep="")
  if(print.progress) cat("Gathering results...\n",sep="")

  object = list(timepoints=timepoints,data=df0,inlafit=r)
  object$linramp$data$y = object$linramp$data$y/rescale.y.factor

  ## compute posterior marginals and posterior marginal means of z^* = t0 (transition onset), dt (transition duration), y0 (initial ramp level), dy (change in ramp level) sigma (amplitude of AR(1) noise) and tau (parameter for correlation of AR(1) noise)
  t0=INLA::inla.emarginal(function(x)x,r$marginals.hyperpar$`Theta1 for idx`); dt=INLA::inla.emarginal(function(x)exp(x),r$marginals.hyperpar$`Theta2 for idx`);y0=INLA::inla.emarginal(function(x)x,r$marginals.hyperpar$`Theta3 for idx`); dy=INLA::inla.emarginal(function(x)x,r$marginals.hyperpar$`Theta4 for idx`);rho = INLA::inla.emarginal(function(x)2/(1+exp(-x))-1,r$marginals.hyperpar$`Theta5 for idx`); sigma = INLA::inla.emarginal(function(x)1/sqrt(exp(x)),r$marginals.hyperpar$`Theta6 for idx`)
  #muvek = linramp(timepoints,t0=t0,dt=dt,y0=y0,dy=dy)

  t0mean = r$summary.hyperpar$mean[1]; t0lower = r$summary.hyperpar$`0.025quant`[1];t0upper = r$summary.hyperpar$`0.975quant`[1]
  #margt0 = inla.tmarginal(function(x)df$age[1]+x/(n-1)*(df$age[n]-df$age[1]),r$marginals.hyperpar$`Theta1 for idx`);
  margt0 = INLA::inla.tmarginal(function(x)df0$x[1]+x/(n-1)*(df0$x[n]-df0$x[1]),r$marginals.hyperpar$`Theta1 for idx`);
  z.t0 = INLA::inla.zmarginal(margt0,silent=TRUE)

  object$linramp$param$t0 = list(marg.t0=margt0,mean=z.t0$mean,sd=z.t0$sd,q0.025=z.t0$quant0.025,q0.5=z.t0$quant0.5,q0.975=z.t0$quant0.975)
  if((abs(r$summary.hyperpar$mean[2])>1000) || (r$summary.hyperpar$sd[2]>1000)){
    margdt=NA
    margdtpos=NA
  }else{
    margdt = INLA::inla.tmarginal(function(x)exp(x)/(n-1)*(df0$x[n]-df0$x[1]),INLA::inla.smarginal(r$marginals.hyperpar$`Theta2 for idx`))
    margdtpos = data.frame(x=-margdt$x,y=margdt$y)
    z.dt = INLA::inla.zmarginal(margdt,silent=TRUE)
    z.dtpos = INLA::inla.zmarginal(margdtpos,silent=TRUE)
    object$linramp$param$dt = list(marg.dt=margdt,mean=z.dt$mean,sd=z.dt$sd,q0.025=z.dt$quant0.025,q0.5=z.dt$quant0.5,q0.975=z.dt$quant0.975)
    object$linramp$param$dtpos = list(marg.dtpos=margdtpos,mean=z.dtpos$mean,sd=z.dtpos$sd,q0.025=z.dtpos$quant0.025,q0.5=z.dtpos$quant0.5,q0.975=z.dtpos$quant0.975)
  }
  margy0 = INLA::inla.tmarginal(function(x)x/rescale.y.factor,
                                INLA::inla.smarginal(r$marginals.hyperpar$`Theta3 for idx`))
  margdy = INLA::inla.tmarginal(function(x)x/rescale.y.factor,
                                INLA::inla.smarginal(r$marginals.hyperpar$`Theta4 for idx`))
  z.y0 = INLA::inla.zmarginal(margy0,silent=TRUE)
  z.dy = INLA::inla.zmarginal(margdy,silent=TRUE)
  object$linramp$param$y0 = list(marg.y0=margy0,mean=z.y0$mean,sd=z.y0$sd,q0.025=z.y0$quant0.025,q0.5=z.y0$quant0.5,q0.975=z.y0$quant0.975)
  object$linramp$param$dy = list(marg.dy=margdy,mean=z.dy$mean,sd=z.dy$sd,q0.025=z.dy$quant0.025,q0.5=z.dy$quant0.5,q0.975=z.dy$quant0.975)

  margsigma = INLA::inla.tmarginal(function(x)1/sqrt(exp(x)),INLA::inla.smarginal(r$marginals.hyperpar$`Theta5 for idx`))
  margtau = INLA::inla.tmarginal(function(x)exp(x),INLA::inla.smarginal(r$marginals.hyperpar$`Theta6 for idx`))
  z.sigma = INLA::inla.zmarginal(margsigma,silent=TRUE)
  z.tau = INLA::inla.zmarginal(margtau,silent=TRUE)
  object$linramp$param$sigma = list(marg.sigma=margsigma,mean=z.sigma$mean,sd=z.sigma$sd,q0.025=z.sigma$quant0.025,q0.5=z.sigma$quant0.5,q0.975=z.sigma$quant0.975)
  object$linramp$param$tau = list(marg.sigma=margtau,mean=z.tau$mean,sd=z.tau$sd,q0.025=z.tau$quant0.025,q0.5=z.tau$quant0.5,q0.975=z.tau$quant0.975)

#nsims = 30000
  if(nsims>0 ) time.startbonussample = Sys.time()

  if(nsims>0 && print.progress ) cat("Simulating ensemble of ", nsims, " samples for t1 = t0 + dt..."," and linear ramp\n",sep="")

  if(nsims>0){

    samps=INLA::inla.hyperpar.sample(nsims,r)

    hpars = matrix(NA,nrow = nsims,ncol=5)
    hpars[,1:2]=samps[,3:4]/rescale.y.factor #y0,dy
    n=length(timepoints)

    hpars[,3] = df0$x[1] + (samps[,1]-1)/(n-1)*(df0$x[n]-df0$x[1]) #t0
    hpars[,4] = exp(samps[,2])/(n-1)*(df0$x[n]-df0$x[1]) #Dt
  }

  if(nsims>0){
    t1sims=numeric(nsims)
    t1mean = mean(t1sims)
    for(i in 1:nsims){
      t01 = hpars[i,3]
      dt1 = hpars[i,4]
      t1sims[i] = t01+dt1

    }

    t1dens = density(t1sims)
    margt1 = cbind(t1dens$x,t1dens$y); colnames(margt1) = c("x","y")
    z.t1 = INLA::inla.zmarginal(margt1,silent=TRUE)
    object$linramp$param$t1 = list(marginal=margt1,samples = t1sims,mean=z.t1$mean,sd=z.t1$sd,q0.025=z.t1$quant0.025,q0.5=z.t1$quant0.5,q0.975=z.t1$quant0.975)
  }
  if(nsims>0){
    vekmat = matrix(NA,nrow=n,ncol=nsims)
    for(i in 1:nsims){
      t01 = hpars[i,3]
      dt1 = hpars[i,4]
      vekmat[,i] = bremla::linramprev(object$data$x,t0=t01,dt=dt1,y0=hpars[i,1],dy=hpars[i,2])
    }
    vek.quant0.025 = numeric(n)
    vek.quant0.5 = numeric(n)
    vek.quant0.975 = numeric(n)
    vek.mean = numeric(n)
    for(iter in 1:n){
      dens = density(vekmat[iter,])
      vek.quant0.025[iter]=INLA::inla.qmarginal(0.05,dens)
      vek.quant0.5[iter]=INLA::inla.qmarginal(0.5,dens)
      vek.mean[iter] = mean(vekmat[iter,])
      vek.quant0.975[iter]=INLA::inla.qmarginal(0.95,dens)
    }

    object$linramp$linrampfit = list(mean = vek.mean,q0.025=vek.quant0.025,q0.5=vek.quant0.5,q0.975=vek.quant0.975)
  }
  #object$time$linramp = list(inla=elapsedinla)
  if(nsims>0 ) object$time$t1_and_ramp = difftime(Sys.time(),time.startbonussample,units="secs")[[1]]
  if((nsims>0 ) && print.progress) cat(" completed in ",object$time$t1_and_ramp," seconds!\n",sep="")

  #time.total = difftime(Sys.time(),time.start,units="secs")[[1]]

  object$linramp$.args = list(nsims=nsims)
  #object$time$linramp=time.total
return(object)
  }

```

```{r linrampplot}
library(ggplot2)
linrampplot = function(object,mainlab="",sublab="", ref=NULL){

  ggd = data.frame(x=object$data$x,y=object$data$y, mean = object$linramp$linrampfit$mean, lower=object$linramp$linrampfit$q0.025,upper=object$linramp$linrampfit$q0.975)

  ggp = ggplot(data=ggd) + geom_line(aes(x=x,y=y),col="gray")+
    geom_ribbon(aes(x=x,ymin=lower,ymax=upper),fill="red",alpha=0.3) + geom_line(aes(x=x,y=mean),col="blue") + theme_bw() + xlab("Depth") + ylab("d18O")+
    labs(title=mainlab, subtitle=sublab) + scale_x_reverse()
  if(!is.null(ref)){
    ggref = data.frame(ref=ref)
    ggp = ggp + geom_vline(data=ggref,aes(xintercept=ref),color="black")
  }

  return(ggp)
}
```

# Fitting linramp model

```{r nextblock}
library(readxl)
library(bremla)
library(numDeriv)

finetunenames = function(strings,tolower=FALSE){
  cstrings = c()
  for(i in 1:length(strings)){
    str = strings[i]
    #if(str_detect(str," ")){
    str = stringcut(str,expr=" ",before=FALSE)
  #}
    if(str_detect(str,"ö")){
      str <- gsub("ö", "o", str)
    }
    cstrings[i] = str
  }

  if(tolower) return(tolower(cstrings))
  return(cstrings)
}

sheetnames = excel_sheets(path="corricdata.xlsx")[-1]
sheetnames = finetunenames(sheetnames)

rowskips = rep(2,length(sheetnames))
rowskips[c(3,8,15,21)] = 3
rowskips[c(11,38)] = 4

#names(df2list)

errorvec = numeric(length(df2list)) #index represents iterator 'i' below. Value represent which event the error occured

margt0_x = c()
margt0_y = c()
recordindices= c()
recordname = c()
eventindices = c()
eventnames = c()

counter = 0
for(i in 1:length(df2list)){

  nam = names(df2list)[i]
  cat("Iteration #",i," of ",length(df2list),": ",nam,"\n\n",sep="")

  locstring = stringcut(nam,"_",before=FALSE)
  groupstring = stringcut(nam,"_",before=TRUE)

  if(locstring %in% sheetnames){
    sheetid = which(tolower(sheetnames) == tolower(locstring))+1
  }else{
    warning(paste0("Could not find correct sheet. ",i,": ",locstring,"\n"))
  }
  groupnames = colnames(read_excel("corricdata.xlsx",sheet=sheetid,n_max=1))[-1]
  groupnames = groupnames[!str_detect(groupnames,"\\.\\.\\.")]
  if(groupstring %in% groupnames){
  }else{
    warning(paste0("Could not find correct group. ",i,": ",locstring," - ",groupstring,"\n"))
  }

  dd = read_excel("corricdata.xlsx",sheet=sheetid,skip=rowskips[sheetid-1],trim_ws=TRUE)
  emptycols = rep(FALSE,length(colnames(dd)))
  emptycolslist = gregexpr("\\.\\.\\.", colnames(dd))
  for(j in 1:length(emptycolslist)){
    if(emptycolslist[[j]][1] == 1)
      emptycols[j] = TRUE
  }
  dd = dd[,!emptycols]
  cnames = colnames(dd)

  agevecs = dd[,str_detect(tolower(cnames),"age_1950")]
  depthvecs = dd[,str_detect(tolower(cnames),"depth")]
  d18Ovecs = dd[,str_detect(tolower(cnames),"d18o")]

  inevents = nrow(df2list[[i]])
  ineventnames = rownames(df2list[[i]])

  for(eventid in 1:inevents){
    counter = counter+1
    ineventname = ineventnames[eventid]
    cat("\n\n",counter,") event #",eventid," of ",inevents,": ",ineventname,"\n\n",sep="")

    eventage = as.numeric(df2list[[i]]$age[eventid])
    colid = which.col(agevecs, eventage)
    if(length(colid)==0){
      errorvec[i] = eventid
      cat("Breaking inner for-loop\n")
      break
      cat("Didnt work!\n")
    }
    if(length(colid)>1){
      warning("Onset found in multiple records, choosing the one where event is closest to middle in terms of indices...\n")
      mincolid = colid[1]
      bestplacement = Inf
      for(ci in 1:length(colid)){
        agve = agevecs[[ colid[ci] ]][!is.na(agevecs[[ colid[ci] ]])]
        deve = depthvecs[[ colid[ci] ]][!is.na(depthvecs[[ colid[ci] ]])]
        if(deve[1] > tail(deve,1)){ #choosing depths that are increasing
        }else{
          #1/length(agve)*sqrt(sum((eventage-agve)^2))

        placement = which(sort(unique(c(agve,eventage)))==eventage) / length(agve)

        offcenter = abs(placement-0.5)

        if(offcenter < bestplacement){
          mincolid = colid[ci]
          bestplacement = offcenter
        }
        }
      }
      if(bestplacement == Inf){
        stop("Could not find suitable record (increasing depths\n")
      }
      colid = mincolid

    }

    depthvec = as.numeric(depthvecs[colid][[1]])
    d18Ovec = as.numeric(d18Ovecs[colid][[1]])
    agevec = as.numeric(agevecs[colid][[1]])

    removeind = c(which(is.na(depthvec)),which(is.na(d18Ovec)),which(is.na(agevec)))
    removeind = unique(removeind)
    if(length(removeind)>0){
      depthvec = depthvec[-removeind]
      agevec = agevec[-removeind]
      d18Ovec = d18Ovec[-removeind]
    }

    if(depthvec[1]>tail(depthvec,1)){
      warning("depth is decreasing. Stopping\n")
      errorvec[i] = colid
      break
    }

    #faultydepths = which(diff(depthvec)<=0)+1
    #faultydepths = which(abs(diff(depthvec))>1)+1
    is.true=TRUE
    while(is.true){
      faultydepths1 = which(diff(depthvec)<=0)+1
      #faultydepths2 = which(abs(diff(depthvec))>1)+1
      #faultydepths = unique(c(faultydepths1,faultydepths2))
      faultydepths = unique(c(faultydepths1))
      if(length(faultydepths)>=1){
        depthvec = depthvec[-faultydepths]
        agevec = agevec[-faultydepths]
        d18Ovec = d18Ovec[-faultydepths]
      }else{
        is.true=FALSE
      }
    }

    #par(mfrow=c(1,2))
    #plot(agevec, d18Ovec,type="l", xlim=rev(range(agevec)))
    #abline(v=eventage)

    ###
    ### insert segmentation ###
    ###

    n = length(agevec)
    eventind = which.index(eventage, agevec)
    upper = min(eventind + round(n/20), n)
    lower = max(eventind - round(n/10), 1)

    interval = rev(seq(from=lower,to=upper,by=1))
    xx = depthvec[interval]
    yy = d18Ovec[interval]
    #plot(xx, yy,type="l",xlim=rev(range(xx)))
    #abline(v=depthvec[eventind],lwd=2,col="blue")

    ### linramp fitting ###

    #remember inversion if needed

    rescale.y.factor = -40/mean(yy)

    results = linrampfunc(xx,yy, steplength=0.005,nsims=5000, opt.params=NULL,imp.fit=TRUE, rescale.y.factor=1, print.progress=TRUE)

    ## temporary placed plot here to give the other plot more time
    plot(agevec, d18Ovec,type="l", xlim=rev(range(agevec)))
    abline(v=eventage)
    #####

    eventname = rownames(df2list[[i]])
    ggp = linrampplot(results,mainlab=ineventname,sublab=nam,ref=depthvec[eventind])
    plot(ggp)
  ### ### ###

    margt0_x = cbind(margt0_x,matrix(results$linramp$param$t0$marg.t0[,1],ncol=1))
    margt0_y = cbind(margt0_y,matrix(results$linramp$param$t0$marg.t0[,2],ncol=1))

    recordindices= c(recordindices, i)
    recordname = c(recordname, nam)
    eventindices = c(eventindices, eventid)
    eventnames = c(eventnames, eventname)

    ggsave(filename=file.path("figs",paste0(counter,"_recordnumber",i,"_event",eventid,"_plot-4800x2400.eps")), device=cairo_ps,width=4800,height=2400,units="px",dpi=500,limitsize=FALSE)

  }

  plot(margt0_x[,1],margt0_y[,1])

}

```