### 2025-01-17 inst/reproduce_results_BA/appendix-stanexample.Rmd
Now supports systematic biases
Systematic biases can now be described using an RW2 process. Also added other stuff, including code to reproduce the results of the most recent paper

# Simulation example in Section 9
This codes generates synthetic layer increments and tie-points and fits the layer-increment and the age-discrepancy models jointly using Markov chain Monte Carlo. Specifically, we use the Stan implementation of the No-U-turn Sampler.

This approach is much more computationally intensive than using the simplification described in the paper along with R-INLA. Hence, the synthetic time series is much shorter.

## Simulate data

```{r}
library(rstan)
library(Matrix)
library(devtools)
library(ggplot2)
set.seed(123)
n = 100

set.seed(123)
y0 = 0
z0 = 0
phi=0.8; sigma=3

depth = z0+seq(from=1,n,length.out=n)
depth2 = depth^2#/z0 #normalize for stability

dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))

truevals = c(sigma,phi,       0.5, 10, 0.01, 20, 0.02, 0, 0.03)
namevals = c("\\sigma","\\phi","b_w","a_1","b_1","a_2","b_2","a_3","b_3")

locs = c(1,25, 75, n)

events=list(locations=locs)
segs = diff(locs); segs[3] = segs[3]+1
a1 = c(rep(1,segs[1]),numeric(segs[2]+segs[3]))
a2 = c(numeric(segs[1]),rep(1,segs[2]),numeric(segs[3]))
a3 = c(numeric(segs[1]+segs[2]),rep(1,segs[3]))
c1 = c(depth[1:(locs[2]-1)],numeric(segs[2]+segs[3]))
c2 = c(numeric(segs[1]),depth[locs[2]:(locs[3]-1)],numeric(segs[3]))
c3 = c(numeric(segs[1]+segs[2]),depth[locs[3]:locs[4]])
dy = dnoise + a1*truevals[4]+a2*truevals[6]+a3*truevals[8]+c1*truevals[5]+
  c2*truevals[7]+c3*truevals[9] + proxy*truevals[3]
truetrend = a1*truevals[4]+a2*truevals[6]+a3*truevals[8]+c1*truevals[5]+
  c2*truevals[7]+c3*truevals[9] + proxy*truevals[3]
#plot(dy)
age = y0+cumsum(dy)
df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth); data=rbind(c(y0,z0,NA,NA),df)
formula=dy~-1+proxy#+depth2
# formula = dy ~ -1 + a1+a2+a3+c1+c2+c3 + proxy
nsims = 10000

# lmr = lm(formula,data=data.frame(dy=dy,proxy=proxy,a1=a1,a2=a2,a3=a3,c1=c1,c2=c2,c3=c3))

res = bremla_prepare(formula, data,nsims, events=events)

design_matrix <- model.matrix(res$.args$formula.ls, res$data)
betatrues = c(truevals[3],truevals[4],truevals[5],truevals[6],truevals[7],truevals[8],truevals[9])

fixedtrue = design_matrix%*%betatrues
#dy = design_matrix%*%betatrues + dnoise

design_matrix <- model.matrix(res$.args$formula.ls, res$data)

{
  ggd = data.frame(true=fixedtrue, dy=dy,depth=depth)
  ggplot(ggd, aes(x=depth)) + theme_bw() + xlab("Depth") + ylab("Layer increments")+
    geom_point(aes(y=dy),col="gray")+
    geom_line(aes(y=fixedtrue),col="black")
}

set.seed(1)
y_truth = cumsum(dy) # just to check where tie-points should be placed

bias = 500*sin((1:n)*3*pi/(2*n))
y_full = y_truth+bias

#which_obs = sort(sample(1:n,n/5))
#which_obs = sort(sample(1:n,n-50))
which_obs = sort(c(10,25,50,75,90))
is_missing = rep(TRUE,n)
is_missing[which_obs] = FALSE
which_missing = which(is_missing)
is_obs = !is_missing
y_full[which_missing]=NA
y_obs = y_full[which_obs]

x_y = seq(from=1,to=10,length.out=n)
stan_data <- list(
  N = n,
  K = ncol(design_matrix),
  X = design_matrix,
  dy = dy,
  N_obs = length(which_obs),
  N_missing = length(which_missing),
  which_missing = which_missing,
  which_obs = which_obs,
  y_obs = y_obs#,
)

```

## Fit model
```{r}

nsims = 4000
warmup = 3000
fit <- stan(
  file = 'AR1plusarima.stan',           # Stan model file
  data = stan_data,                  # Data list for Stan
  iter = nsims,                       # Number of iterations
  warmup = warmup,
  chains = 4,                        # Number of chains
  seed = 123                         # Seed for reproducibility
) 
samples <- extract(fit)

dim(samples$y)
ycolm = colMeans(samples$y)
ycolq = colQuantiles(samples$y, probs=c(0.025,0.975))
Xmucolm = colMeans(samples$X_mu)
Xmucolq = colQuantiles(samples$X_mu, probs=c(0.025,0.975))

{
  ggd1 = data.frame(x=depth,mean=Xmucolm, lower=Xmucolq[,1], upper=Xmucolq[,2], obs = dy, truetrend = truetrend)
  ggp1 = ggplot(data=ggd1, aes(x=x)) + theme_bw() + xlab("Depth") + ylab("Layers per unit of depth")+
    ggtitle("(a) Fitted layer increments") +
    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
    geom_point(aes(y=obs),col="black")+
    #geom_line(aes(y=obs),col="gray")+
    geom_line(aes(y=truetrend),col="blue")+
    geom_line(aes(y=mean))
}

{
  cumsumXmu = matrix(NA,ncol=ncol(samples$X_mu),nrow=nrow(samples$X_mu))
  for(i in 1:nrow(samples$X_mu)){
    cumsumXmu[i,] = cumsum(samples$X_mu[i,])
  }
  cumxm = colSums(cumsumXmu)
  cumxquant = colQuantiles(cumsumXmu, probs=c(0.025,0.975))

  #plot(y_noise)
  #y_full = y_truth+bias
  y_og = cumsum(dy)
  #ggd2 = data.frame(x=depth,mean=ycolm-y_og,lower=ycolq[,1]-y_og,upper=ycolq[,2]-y_og, bias=bias)
  ggd2 = data.frame(x=depth,mean=ycolm,lower=ycolq[,1],upper=ycolq[,2], bias=bias+cumsum(dy),
                    xmean = cumxm, xlower=cumxquant[,1],xupper=cumxquant[,2])
  ggp2 = ggplot(data=ggd2,aes(x=x)) + theme_bw()+xlab("Depth")+ylab("Estimated age") +
    ggtitle("(b) Synchronized chronologies") +
    geom_ribbon(aes(ymin=xlower,ymax=xupper),fill="black", alpha=0.2)+
    geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3)+
    geom_line(aes(y=bias),col="blue")+
    geom_line(aes(y=mean))+
    #geom_point(data=data.frame(x=depth[which_obs],y=y_obs-y_og[which_obs]),aes(x=x,y=y),col="black")
    geom_point(data=data.frame(x=depth[which_obs],y=y_obs),aes(x=x,y=y),col="black")

  print(ggp2)
}
library(ggpubr)
ggboth1 = ggarrange(ggp1,ggp2,nrow=2)
#ggboth2 = ggarrange(ggp1,ggp3,nrow=1)

#ggpall = ggarrange(ggp11,ggp222,nrow=2)
ggsave(paste0("MCMCexample-3000x2000.eps"),ggboth1,device=cairo_ps, width=3000, height=2200, units="px")

#ggsave("MCMCexample2-4800x2000.eps", device=cairo_ps,plot=ggboth2,width=4800,height=2000, units="px",dpi=500,limitsize=FALSE)

traceplot(fit)
```