### 2025-01-17 inst/reproduce_results_BA/results.Rmd

# Code for Myrvoll-Nilsen et al.

This file includes the data and code needed to reproduce the results for

Myrvoll-Nilsen, E., Riechers, K. & Boers, N. (202x). Synchronization of layer-counted paleoclimatic proxy archives.

## Illustration example

The data used to generate the illustrative examples in the paper is as follows.

```{r illustrationexample}
# rm(list=ls())
require(stats)
library(ggpubr)
library(ggplot2)
set.seed(123)
n=800
y0 = 0
z0 = 0
phi=0.8; sigma=15
depth = z0+seq(from=1,to=n,length.out=n)

covariate = arima.sim(model=list(ar=c(0.99)),n=n)
noise = sigma*arima.sim(model=list(ar=c(phi)),n=n,sd=sqrt(1-phi^2))

dy = 10 + depth*5/n+covariate+noise
y = y0+cumsum(dy)

data=data.frame(age=c(y0,y),depth=c(z0,depth),dy=c(0,as.numeric(dy)),covariate=c(0,covariate))
#data=rbind(c(y0,z0,0,0),df)
formula=dy~ 1 + depth + covariate 
nsims = 10000

# Unsynchronized
res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
             control.fit=list(noise="ar1"),
             x.label="Depth",
             y.label="Ensemble age - observed age",
             control.sim = list(synchronized=FALSE,
                                nsims=nsims))

library(ggplot2)
orichron = res$data$age
ggd1 = data.frame(depth=res$data$depth,
                 mean = res$simulation$summary$mean-orichron,
                 lower = res$simulation$summary$lower-orichron,
                 upper = res$simulation$summary$upper-orichron)
ggd1_2 = data.frame(depth=res$data$depth,
                 mean = res$simulation$summary$mean,
                 lower = res$simulation$summary$lower,
                 upper = res$simulation$summary$upper)
ggpunc = ggplot(data=ggd1,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,linewidth=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)
ggpunc2 = ggplot(data=ggd1_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,linewidth=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)
# plot(res)
#ggsave("illustrative-unsync.eps",device=cairo_ps,width=4800,height=2000,units="px")

# Fixed tie-points
ntie = 3
tiesims = matrix(NA,nrow=nsims,ncol=ntie)
tiedepths = c(100,400,600)
#tieage = c(450,3000,8500)
tieage = c(450,2000,10000)

for(i in 1:ntie){
  tiesims[,i] = rep(tieage[i],nsims)
}

res2 = tiepointsimmer(res,synchronization = list(locations=tiedepths,locations_unit="depth",
                       samples = tiesims)
               )

res2 = bremla_synchronized_simulation(res2,control.sim=list(synchronized=TRUE))

ggdfix = data.frame(depth=res2$data$depth,
                 mean = res2$simulation$summary_sync$mean-orichron,
                 lower = res2$simulation$summary_sync$lower-orichron,
                 upper = res2$simulation$summary_sync$upper-orichron)
ggd_2 = data.frame(depth=res2$data$depth,
                 mean = res2$simulation$summary_sync$mean,
                 lower = res2$simulation$summary_sync$lower,
                 upper = res2$simulation$summary_sync$upper)

oritie = orichron[tiedepths]
ggd2 = data.frame(depth=res2$tie_points$locations,
                  mean=res2$simulation$summary_sync$mean[tiedepths]-oritie)
ggd2_2 = data.frame(depth=res2$tie_points$locations,
                  mean=res2$simulation$summary_sync$mean[tiedepths])

ggpfix = ggplot(data=ggdfix,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(data=ggd1,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.6)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_point(data=ggd2,mapping=aes(x=depth,y=mean),col="black",size=1)
ggpfix2 = ggplot(data=ggd_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(data=ggd1_2,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.6)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_point(data=ggd2_2,mapping=aes(x=depth,y=mean),col="black",size=1)

#plot(res)
#ggsave("illustrative-fixedsync.eps",device=cairo_ps,width=4800,height=2000,units="px")

#tiemeans = c(450,3000,8500)
tiemeans = c(450,2000,10000)
tiesds = c(400,500,800)
res3 = tiepointsimmer(res,synchronization = list(locations=tiedepths,
                                                locations_unit="depth",
                                                method="normal",
                                                params=list(mean=tiemeans,
                                                            sd=tiesds))
               )
res3 = bremla_synchronized_simulation(res3,control.sim=list(synchronized=TRUE))

ggd = data.frame(depth=res3$data$depth,
                 mean = res3$simulation$summary_sync$mean-orichron,
                 lower = res3$simulation$summary_sync$lower-orichron,
                 upper = res3$simulation$summary_sync$upper-orichron)
ggd_2 = data.frame(depth=res3$data$depth,
                 mean = res3$simulation$summary_sync$mean,
                 lower = res3$simulation$summary_sync$lower,
                 upper = res3$simulation$summary_sync$upper)

oritie = orichron[tiedepths]
ggd2 = data.frame(depth=res3$tie_points$locations,
                  lower=res3$simulation$summary_sync$lower[tiedepths]-oritie,
                  upper=res3$simulation$summary_sync$upper[tiedepths]-oritie)
ggd2_2 = data.frame(depth=res3$tie_points$locations,
                  lower=res3$simulation$summary_sync$lower[tiedepths],
                  upper=res3$simulation$summary_sync$upper[tiedepths])

ggprand = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(data=ggd1,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(data=ggd,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_segment(data=ggd2,mapping=aes(x=depth,xend=depth,y=lower,yend=upper), 
               col="black",size=0.7)
ggprand2 = ggplot(data=ggd_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(data=ggd1_2,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_segment(data=ggd2_2,mapping=aes(x=depth,xend=depth,y=lower,yend=upper), 
               col="black",size=0.7)

#plot(res3)
ylim = range(res$simulation$summary$lower-orichron,res$simulation$summary$upper-orichron,
             res3$simulation$summary_sync$lower-orichron,res3$simulation$summary_sync$upper-orichron)
ggpunc = ggpunc + ylim(ylim) + ggtitle("(a) Unsynchronous chronologies")
ggpfix = ggpfix + ylim(ylim)+ ggtitle("(b) Synchronous chronologies: fixed tie-points")
ggprand = ggprand + ylim(ylim)+ ggtitle("(c) Synchronous chronologies: uncertain tie-points")
ggplotlist0 = c(); ggplotlist0[[1]] = ggpunc; ggplotlist0[[2]] = ggpfix; ggplotlist0[[3]] = ggprand
ggplotlist00 = ggarrange(plotlist=ggplotlist0,nrow=3,ncol=1)

ylim = range(res$simulation$summary$lower,res$simulation$summary$upper,
             res3$simulation$summary_sync$lower,res3$simulation$summary_sync$upper)
ggpunc2 = ggpunc2 + ylim(ylim) + ggtitle("(a) Unsynchronous chronologies")+ylab("Ensemble age")
ggpfix2 = ggpfix2 + ylim(ylim)+ ggtitle("(b) Synchronous chronologies: fixed tie-points")+ylab("Ensemble age")
ggprand2 = ggprand2 + ylim(ylim)+ ggtitle("(c) Synchronous chronologies: uncertain tie-points")+ylab("Ensemble age")
ggplotlist1 = c(); ggplotlist1[[1]] = ggpunc2; ggplotlist1[[2]] = ggpfix2; ggplotlist1[[3]] = ggprand2
ggplotarrange = ggarrange(plotlist=ggplotlist1,nrow=3,ncol=1)
plot(ggplotarrange)

# ggsave("illustrative-all-3000x3400.eps",device=cairo_ps,width=3000,height=3400,units="px",dpi=500,limitsize=FALSE)

#ggprand2 = ggprand2 + ylim(ylim)+ ggtitle("(c) Synchronous chronologies: uncertain tie-points")+ylab("Ensemble age")
# 
# ggprand22 =  ggprand2 + geom_ribbon(data=ggd1_2,mapping=aes(x=depth, ymin=lower,ymax=upper),fill="black",alpha=0.2)
# ggpfix22 =  ggpfix2 + geom_ribbon(data=ggd1_2,mapping=aes(x=depth, ymin=lower,ymax=upper),fill="black",alpha=0.2)

ggplotlist1 = c(); ggplotlist1[[1]] = ggpunc2; ggplotlist1[[2]] = ggpfix2; ggplotlist1[[3]] = ggprand2
ggplotlist2 = c(); ggplotlist2[[1]] = ggpunc2; ggplotlist2[[2]] = ggpfix22; ggplotlist2[[3]] = ggprand22
ggplotarrange = ggarrange(plotlist=ggplotlist1,nrow=3,ncol=1)
plot(ggplotarrange)
ggplotarrange2 = ggarrange(plotlist=ggplotlist2,nrow=3,ncol=1)

print(ggplotarrange2)

#ggsave("illustrative-all2-3000x3400.eps",device=cairo_ps,width=3000,height=3400,units="px",dpi=500,limitsize=FALSE)

```

# Simulation example

```{r}
require(stats)
library(ggplot2)
set.seed(123)
n=1000
y0 = 0
z0 = 0
phi=0.8; sigma=3

depth = z0+seq(from=1,1000,length.out=1000)
depth2 = depth^2#/z0 #normalize for stability

dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))

truevals = c(sigma,phi,       0.5, 10, 0.01, 20, 0.02, 0, 0.03)
namevals = c("\\sigma","\\phi","b_w","a_1","b_1","a_2","b_2","a_3","b_3")

events=list(locations=c(1,250,750,1000))
a1 = c(rep(1,249),numeric(751))
a2 = c(numeric(249),rep(1,500),numeric(251))
a3 = c(numeric(749),rep(1,251))
c1 = c(depth[1:249],numeric(751))
c2 = c(numeric(249),depth[250:749],numeric(251))
c3 = c(numeric(749),depth[750:1000])
dy = dnoise + a1*truevals[4]+a2*truevals[6]+a3*truevals[8]+c1*truevals[5]+
  c2*truevals[7]+c3*truevals[9] + proxy*truevals[3]
plot(dy)
age = y0+cumsum(dy)
df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth); data=rbind(c(y0,z0,NA,NA),df)
formula=dy~-1+proxy#+depth2
nsims = 1000

control.sim = list(ncores=1, nsims=nsims, synchronized=TRUE)
# Specify tie-point distribution
control.fit=list(method="INLA", noise="ar1")
synchronization = list(locations = c(1, 200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(0, 2200,6800,19000),sd=c(0, 10,50,60)), 
                       agedisc=list(model="rw2",
                                    options=list(restart.fromlast=FALSE,
                                                 inla.options=list(num.threads=1,
                                                                   verbose=TRUE,
                                                                   control.compute=list(openmp.strategy="huge")))))

synchronization = list(locations = c(1, 200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(0, 2200,6800,19000),sd=c(0, 10,50,60)), 
                       agedisc=list(model="rw2"))

```

```{r}

## set up model
res = bremla_prepare(formula=formula, data=data, nsims=nsims, events=events,
               control.fit=control.fit,
               control.sim=control.sim,
               synchronization=synchronization)

## fit model
res = bremla_modelfitter(res)

## sample unsynchronized chronologies

res = bremla_chronology_simulation(res)

## generate tie-points

res = tiepointsimmer(res)

## sample synchronized chronologies
res = bremla_synchronized_simulation(res, print.progress=TRUE)

## To do all these steps in one call use:
# res = bremla(formula=formula, data=data, nsims=nsims, events=events,
#                control.fit=control.fit,
#                control.sim=control.sim,
#                synchronization=synchronization,
#                print.progress=TRUE)

```

# Real data example: NGRIP/GICC05

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). 
We also use tie-points described by probability distributions obtained from Adolphi et al. (2018) and Muscheler et al. (2020). Both the NGRIP/GICC05 data and the Adolphi/Muscheler tie-points are included in the package and can be imported as follows.

```{r}
library(INLA)
set.seed(1991)
require(stringr)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
d18O = NGRIP_5cm$d18O
proxy=d18O
formula = dy~-1+depth2
depth2 = depth^2/depth[1]^2 #normalize for stability

# set up data set with n+1 rows. Must include 'age', 'depth' and all covariates in 'formula'. In the first row, only 'y0' and 'z0' are collected:
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)

nsims=1000 #number of chronologies to be sampled

# specify the events which separates piecewise predictor in layer increment model. See '?events.default' for more details:
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

# Simulation options. Specify 'synchronization=TRUE' so it doesnt only create unsynchronized samples. See '?control.sim.default' for more details
control.sim=list(synchronized = TRUE)

# Fit options. These specify options for the model and fitting procedure in the layer-increment model. We use AR(1) noise and the INLA method. See '?control.fit.default' for more details
control.fit=list(method="inla", noise="ar1")

# Options for synchronization approach. Here, tie-point samples, or model specifications must be included. locations should also be included. To use the Adolphi tie-points, specify 'method="adolphi"' and use 'locations' to specify which tie-points to be included as demonstrated below:
synchronization=list(method="adolphi", #Choose which Adolphi tie-point to use (the first one takes place after the Holocene and is therefore omitted here)
                     locations=c(FALSE,TRUE,TRUE,TRUE,TRUE)
                     )

```

## Fit models with bremla

The bremla package will set up the bremla object with all options specified in 'events', 'control.fit', 'control.sim' and 'synchronization', using the 'bremla_prepare' function. It also contains functions to perform least squares to find good initial values for an INLA fit, and simulate 
It will also set up the bremla object with all

### Assuming no systematic biases
If we assume there are no systematic biases, then the tie-points can be considered as realizations of the cumulative of the inferred layer increment model. If this model is Gaussian we can do this very efficiently

```{r}

 object0 = bremla(formula=formula, data=data, nsims=nsims, events=events,
                control.fit=control.fit,
                control.sim=control.sim,
                synchronization=synchronization)

```

#### Plot results

```{r}

ggdreal0 = data.frame(depth=object0$data$depth, age=object0$data$age, agemean= object0$simulation$summary$mean, agelower = object0$simulation$summary$lower, ageupper=object0$simulation$summary$upper, sagemean=object0$simulation$summary_sync$mean, 
                 sagelower=object0$simulation$summary_sync$lower, sageupper=object0$simulation$summary_sync$upper,
                 agediffmean=object0$simulation$summary$mean-object0$data$age,
                 agedifflower=object0$simulation$summary$lower-object0$data$age, 
                 agediffupper=object0$simulation$summary$upper-object0$data$age,
                 sagediffmean=object0$simulation$summary_sync$mean-object0$data$age,
                 sagedifflower=object0$simulation$summary_sync$lower-object0$data$age,
                 sagediffupper=object0$simulation$summary_sync$upper-object0$data$age)

#library(matrixStats)
tiequants = colQuantiles(object$tie_points$samples, probs=c(0.025,0.975))
ggdtiereal = data.frame(locations=object$data$depth[object$tie_points$locations_indexes],
                        lower = tiequants[,1], upper = tiequants[,2],
                        dlower = tiequants[,1] - object$data$age[object$tie_points$locations_indexes], dupper = tiequants[,2] - object$data$age[object$tie_points$locations_indexes])

library(ggplot2)

ggpreal0 = ggplot(data=ggdreal0,aes(x=depth)) + theme_bw()+
  geom_ribbon(aes(ymin=agedifflower, ymax=agediffupper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=sagedifflower,ymax=sagediffupper), color="red", fill="red", alpha=0.3)+
  geom_line(aes(y=sagediffmean),color="blue") +
  geom_linerange(data=ggdtiereal,aes(x=locations,ymin=dlower,ymax=dupper),color="black")+
  xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)")

```

### Using RW2 bias
Here we assume there could be systematic biases, and model the resulting discrepancy between the number of inferred layers and the true age, as represented by the tie-points, using a 2nd order random walk model

```{r}

## set up model

synchronization$agedisc = list(model="rw2") #choose the 2nd order random walk for the age discrepancy model

object = bremla_prepare(formula=formula, data=data, nsims=nsims, events=events,
               control.fit=control.fit,
               control.sim=control.sim,
               synchronization=synchronization)

## fit model
object = bremla_modelfitter(object)

## sample unsynchronized chronologies

object = bremla_chronology_simulation(object)

## generate tie-points

object = tiepointsimmer(object)

## sample synchronized chronologies
object = bremla_synchronized_simulation(object, print.progress=TRUE)

## To do all these steps at the same time use
# object = bremla(formula=formula, data=data, nsims=nsims, events=events,
#                control.fit=control.fit,
#                control.sim=control.sim,
#                synchronization=synchronization,
#                print.progress=TRUE)

```

#### Plot results

```{r}
ggdreal = data.frame(depth=object$data$depth, age=object$data$age, agemean= object$simulation$summary$mean, agelower = object$simulation$summary$lower, ageupper=object$simulation$summary$upper, sagemean=object$simulation$summary_sync$mean, 
                 sagelower=object$simulation$summary_sync$lower, sageupper=object$simulation$summary_sync$upper,
                 agediffmean=object$simulation$summary$mean-object$data$age,
                 agedifflower=object$simulation$summary$lower-object$data$age, 
                 agediffupper=object$simulation$summary$upper-object$data$age,
                 sagediffmean=object$simulation$summary_sync$mean-object$data$age,
                 sagedifflower=object$simulation$summary_sync$lower-object$data$age,
                 sagediffupper=object$simulation$summary_sync$upper-object$data$age)

#library(matrixStats)
tiequants = colQuantiles(object$tie_points$samples, probs=c(0.025,0.975))
ggdtiereal = data.frame(locations=object$data$depth[object$tie_points$locations_indexes],
                        lower = tiequants[,1], upper = tiequants[,2],
                        dlower = tiequants[,1] - object$data$age[object$tie_points$locations_indexes], dupper = tiequants[,2] - object$data$age[object$tie_points$locations_indexes])

library(ggplot2)
ggpreal1 = ggplot(data=ggdreal,aes(x=depth)) + theme_bw()+
  geom_ribbon(aes(ymin=agelower, ymax=ageupper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=sagelower,ymax=sageupper), color="red", fill="red", alpha=0.3)+
  geom_line(aes(y=sagemean),color="blue") +
  geom_linerange(data=ggdtiereal,aes(x=locations,ymin=lower,ymax=upper),color="black")+
  xlab("Depth (m)") + ylab("Age (yr b2k)")+
  ggtitle("(a) Absolute time scale")

ggpreal2 = ggplot(data=ggdreal,aes(x=depth)) + theme_bw()+
  geom_ribbon(aes(ymin=agedifflower, ymax=agediffupper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=sagedifflower,ymax=sagediffupper), color="red", fill="red", alpha=0.3)+
  geom_line(aes(y=sagediffmean),color="blue") +
  geom_linerange(data=ggdtiereal,aes(x=locations,ymin=dlower,ymax=dupper),color="black")+
  xlab("Depth (m)") + ylab("Age deviation (years)")+
  ggtitle("(b) Age deviation from GICC05")

```